### 2019-03-10 DesignDocs/DesignDocWeights.tex
reversing weights work in glmm.R, updating weights design doc

\documentclass{article}

 \usepackage{url} 
\usepackage{amsthm,amsmath,amssymb,indentfirst,float}
\usepackage{verbatim}
\usepackage[sort,longnamesfirst]{natbib}
\newcommand{\pcite}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\ncite}[1]{\citeauthor{#1}, \citeyear{#1}}
\DeclareMathOperator{\logit}{logit}
    \DeclareMathOperator{\var}{Var}
   %  \DeclareMathOperator{\det}{det}
     \DeclareMathOperator{\diag}{diag}

\usepackage{geometry}
%\geometry{hmargin=1.025in,vmargin={1.25in,2.5in},nohead,footskip=0.5in} 
%\geometry{hmargin=1.025in,vmargin={1.25in,0.75in},nohead,footskip=0.5in} 
%\geometry{hmargin=2.5cm,vmargin={2.5cm,2.5cm},nohead,footskip=0.5in}

\renewcommand{\baselinestretch}{1.25}

\usepackage{amsbsy,amsmath,amsthm,amssymb,graphicx}

\setlength{\baselineskip}{0.3in} \setlength{\parskip}{.05in}

\newcommand{\cvgindist}{\overset{\text{d}}{\longrightarrow}}
\DeclareMathOperator{\PR}{Pr} 
\DeclareMathOperator{\cov}{Cov}

\newcommand{\sX}{{\mathsf X}}
\newcommand{\tQ}{\tilde Q}
\newcommand{\cU}{{\cal U}}
\newcommand{\cX}{{\cal X}}
\newcommand{\tbeta}{\tilde{\beta}}
\newcommand{\tlambda}{\tilde{\lambda}}
\newcommand{\txi}{\tilde{\xi}}

\title{Design Document for Relevance-Weighting: R Package glmm}

\author{Sydney Benson}

\begin{document}
\maketitle{}

\begin{abstract}
This design document will give an overview of the changes made to the R package \texttt{glmm} with respect to the relevance-weighted likelihood method mentioned in Wang (2001). We use relevance-weighting to better reflect the real-world occurrence of more or less informative observations.
\end{abstract}

\section{Introduction}
This project is meant to enable the user of the \texttt{glmm} function in the \texttt{glmm} R package to include an optional relevance-weighting scheme. A common assumption of linear models is that each observation in a data set is equally informative and trustworthy; however, in real-world data sets, this is rarely the case. Thus, the optional relevance-weighting scheme will allow the user to place a heavier weight on the more informative and/or trustworthy observations in their data set so that those data points that are less informative affect the model to a lesser degree.

\section{The Process}
The weighting scheme for the observations can be implemented in the \texttt{objfun.R} function. First, the function will need to establish whether the user has supplied a proper weighting scheme. Next, the weighting scheme will need to be applied to the respective elements of the log-likelihood. After defining the weighting vector, the remainder of this section will illustrate how this weighting scheme will be applied to the log-likelihood values.

\subsection{The Weighting Vector}
This vector, called $\Lambda$, must be a vector with the same length as the response vector and must contain all positive values.

\subsection{Weighted Log-Likelihood}
From Hu and Zidek's (1997) method, mentioned in Wang (2001), we know the relevance-weighted likelihood (REWL) is

\begin{align}
\text{REWL}(\theta) = \prod_{i=1}^{n} {f(x_i;\theta)}^{\lambda_i}
\end{align}

\noindent where $f(x_i;\theta)$ is the probability distribution for $x_i$ and $\lambda_i$ is the weight for $x_i$. Then, the relevance-weighted log-likelihood (RWLL) is

\begin{align}
\text{RWLL}(\theta) &= \log\left(\prod_{i=1}^{n} {f(x_i;\theta)}^{\lambda_i}\right) \\
&= \sum_{i=1}^{n} \log\left({f(x_i;\theta)}^{\lambda_i}\right) \\
&= \sum_{i=1}^{n} \lambda_i \log\left(f(x_i;\theta)\right)
\end{align}

Then, knowing that $l_m(\theta | y) = \log\left(\dfrac{1}{m}\sum_{i=1}^{r}j_ie^{v_i}\right)$ where $m$ is the number of rows of the $u$ matrix, $j_i$ is the number of rows of the chunk of the $u$ matrix being processed by the $i$th core and $v_i$ is the value obtained by each core, as described in the design document for parallel computing, we have

\begin{align}
\text{RWLL}(\theta) &= \sum_{i=1}^{n} \lambda_i \log\left(\dfrac{1}{m}\sum_{i=1}^{r}j_ie^{v_i}\right) \\
&= a + \sum_{i=1}^{n} \lambda_i \log\left(\dfrac{1}{m}\sum_{i=1}^{r}j_ie^{v_i - a}\right)
\end{align}

\noindent where $a = \max(v_i)$.

\end{document}