### 2019-04-11 DesignDocs/DesignDocWeights.tex
finished weights design doc

\documentclass{article}

 \usepackage{url} 
\usepackage{amsthm,amsmath,amssymb,indentfirst,float}
\usepackage{verbatim}
\usepackage[sort,longnamesfirst]{natbib}
\newcommand{\pcite}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\ncite}[1]{\citeauthor{#1}, \citeyear{#1}}
\DeclareMathOperator{\logit}{logit}
    \DeclareMathOperator{\var}{Var}
   %  \DeclareMathOperator{\det}{det}
     \DeclareMathOperator{\diag}{diag}

\usepackage{geometry}
%\geometry{hmargin=1.025in,vmargin={1.25in,2.5in},nohead,footskip=0.5in} 
%\geometry{hmargin=1.025in,vmargin={1.25in,0.75in},nohead,footskip=0.5in} 
%\geometry{hmargin=2.5cm,vmargin={2.5cm,2.5cm},nohead,footskip=0.5in}

\renewcommand{\baselinestretch}{1.25}

\usepackage{amsbsy,amsmath,amsthm,amssymb,graphicx}

\setlength{\baselineskip}{0.3in} \setlength{\parskip}{.05in}

\newcommand{\cvgindist}{\overset{\text{d}}{\longrightarrow}}
\DeclareMathOperator{\PR}{Pr} 
\DeclareMathOperator{\cov}{Cov}

\newcommand{\sX}{{\mathsf X}}
\newcommand{\tQ}{\tilde Q}
\newcommand{\cU}{{\cal U}}
\newcommand{\cX}{{\cal X}}
\newcommand{\tbeta}{\tilde{\beta}}
\newcommand{\tlambda}{\tilde{\lambda}}
\newcommand{\txi}{\tilde{\xi}}

\title{Design Document for Relevance-Weighting: R Package glmm}

\author{Sydney Benson}

\begin{document}
\maketitle{}

\begin{abstract}
This design document will give an overview of the changes made to the R package \texttt{glmm} with respect to a relevance-weighted likelihood method. We use relevance-weighting to better reflect the real-world occurrence of more or less informative observations.
\end{abstract}

\section{Introduction}
This project is meant to enable the user of the \texttt{glmm} function in the \texttt{glmm} R package to include an optional relevance-weighting scheme. A common assumption of linear models is that each observation in a data set is equally informative and trustworthy; however, in real-world data sets, this is rarely the case. Thus, the optional relevance-weighting scheme will allow the user to place a heavier weight on the more informative and/or trustworthy observations in their data set so that those data points that are less informative affect the model to a lesser degree.

\section{The Process}
First, the function will need to establish whether the user has supplied a proper weighting scheme. Next, the weighting scheme will need to be applied in the \texttt{el.C} function. After defining the weighting vector, the remainder of this section will illustrate how this weighting scheme will be applied.

\subsection{The Weighting Vector}
The weights must be in the form of a vector and the vector must be the same length as the response vector, $y$. All weights must be numeric and non-negative. There can be no missing weights.

\subsection{The Relevance-Weighted Log Density}
Following Hu and Zidek (1997), the relevance-weighted likelihood is defined as

\begin{align}
\text{REWL}(\theta) = \prod_{i=1}^n f(y_i | \theta)^{\lambda_i}
\end{align}

\noindent where $n$ is the length of the response vector, $y$. Since we define $\log f_\theta(y|u_k)$ as $\sum_i y_i\eta_i - c(\eta_i)$, where $\eta = X\beta + Zu$, $\log f_\theta(y_i | u_k) = y_i\eta_i - c(\eta_i)$ and $f_\theta(y_i|u_k) = \exp \left(  y_i\eta_i - c(\eta_i) \right)$. So, the relevance-weighted log density becomes

\begin{align}
\text{RWLD}(\theta) &= \prod_{i=1}^n \left[ \exp \left(  y_i\eta_i - c(\eta_i) \right) \right] ^{\lambda_i}  \\
&= \prod_{i=1}^n \exp \left[ \lambda_i \left(  y_i\eta_i - c(\eta_i) \right) \right]
\end{align}

\noindent Then,

\begin{align}
\log \text{RWLD}(\theta) &= \log \prod_{i=1}^n \exp \left[ \lambda_i \left(  y_i\eta_i - c(\eta_i) \right) \right] \\
&= \sum_{i=1}^n \log \exp \left[ \lambda_i \left(  y_i\eta_i - c(\eta_i) \right) \right] \\
&= \sum_{i=1}^n \lambda_i \left(  y_i\eta_i - c(\eta_i) \right)
\end{align}

\subsection{The First Derivative}

Remember that the derivative of the log density of the data with respect to one component, $\eta_j$, is

\begin{align}
\dfrac{\partial}{\partial \eta_j} \log f_\theta (y|u) = y_j - c'(\eta_j).
\end{align}

\noindent Consequently,

\begin{align}
\dfrac{\partial}{\partial \eta_j} \log \text{RWLD}(\theta) = \lambda_j \left[ y_j - c'(\eta_j) \right].
\end{align}

\noindent and the derivative of the component $\eta_j$ with respect to one of the fixed effect predictors, $\beta_j$, is

\begin{align}
\dfrac{\partial \eta_j}{\partial \beta_l} = X_{jl}
\end{align}

\noindent Then, we can use the chain rule to find that

\begin{align}
\dfrac{\partial}{\partial \beta_l} \log \text{RWLD}(\theta) &= \lambda \ X \left[ y- c'(\eta) \right]
\end{align}

\subsection{The Second Derivative}

Similar to the first derivative, we find that the second derivative of the relevance-weighted log density of the data is

\begin{align}
\dfrac{\partial^2}{\partial \beta_l^2} \log \text{RWLD}(\theta) &= \lambda \ X' \left[- c''(\eta) \right] X
\end{align}

\end{document}